---
title: "helpers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{helpers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rotsee)
```

## facilitating input pre-processing
The rotsee package is built on the mindset that it should be universally applicable.
Hence, location specific calculations such as bulk density determination via pedotransfer
functions are omitted from the core calculation code. While this makes package usage more
flexible, it does put additional requirements on the package user. To facilitate this,
several helper functions have been added. These functions are outlined below, together with
a short description of their origin and an example of using the functions.


## Dry soil bulk density
Dry soil bulk density is used to calculate soil carbon (C) content (e.g. g C/kg soil) to a C stock 
(e.g. kg C/ha) and vice versa. Calculations are always conducted based on C stocks, and hence
providing an accurate dry bulk density is important particularly when C content is given, or
when outputs are requested in C content.

In case the dry soil bulk density has not been measured, it can be calculated using 
\link{rc_calculate_bd}. This function estimates soil bulk density from soil clay-
and C content based on the pedotransfer functions listed in CBAV, 2023. Provided input is returned
together with the column A_DENSITY_SA, listing soil bulk density in g/cm^3.

## Calculate crop C inputs
To remove location specific differences in crop quality, harvest efficiency, etc., rotsee
requests crop input to directly list the total 

## Evapotranspiration factors
TO DO once facil-refact has been merged

## Extend crop rotation plan

## Extend amendment input

## references
